*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-21 Code/Manuscript_O2.Rmd
PUR results edited

Protein samples were thawed and placed in 2 mL FastPrep®-24 bead lysing matrix D vials (MP Biomedicals, Irvine, CA, USA; SKU 116913050) with 200 uL of 1X protein extraction buffer (0.0281 mol L^−1^ TRIS base buffer, 0.0211 mol L^−1^ TRIS HCl buffer, 0.0735 mol L^−1^ lithium dodecyl sulfate, 1.08 mol L^−1^ glycerol, 0.5 mmol L^−1^ ethylenediaminetetraacetic acid, 0.1 mg·mL−1 PefaBloc SC (AEBSF) protease inhibitor) [@brownFluxCapacitiesAcclimation2008]  XXX note we used a low tris PEB recipe not sure where it came from (citation??) XXX. Cells were lysed for 3 cycles of 60 seconds at 6.5m s-1 with 60 second intervals on ice between cycles.  The lysing vials were centrifuged at 16100g using XXXXXX centrifuge ( reference here    ) for 5 minutes, the supernatant was transferred to a 600 uL microcentrifuge tube and centrifuged again at 16100g for 3 minutes. 40 µL of the supernatant were subsampled into labeled vials and stored at -80 °C until analysis. The bicinchoninic acid (BCA) assay was used for total protein determination using a calibration curve of bovine gamma globulin (BGG; Bio-Rad, Hercules, CA, USA) standard ranging from 0 to 1.0 mg mL^-1^ range.

All samples and appropriate standards were prepared as described in [@brownFluxCapacitiesAcclimation2008]. Standards were obtained from Agrisera (Vännäs, Sweden; FtsH AS11 1789S; PsbA AS01 016S; PetC AS08 330S; PsaC AS04 042S; RbcL AS01 017S).  Between 0.1 to 4.0 µg of total protein was loaded onto 15 well Invitrogen Bolt 4-12% acrylamide Bis-Tris gels (ThermoFisher Scientific, Waltham, MA, USA). All gels were run for 25 minutes at 200V using Invitrogen Bolt Mini Gel tanks and Invitrogen Bolt MES SDS running buffer (ThermoFisher Scientific, Waltham, MA, USA) and transferred for 60 minutes at 20V onto a PVDF membrane (Bio-Rad, Hercules, CA, USA; SKU 1620177). Blots were blocked overnight in 2% ECL blocking solution as described by @brownFluxCapacitiesAcclimation2008. The following morning, each blot was incubated in their appropriate antibody obtained from Agrisera (Vännäs, Sweden; FtsH  1:5,000 dilution, AS11 1789; PsbA  1:10000 dilution, AS05 084; PetC 1:XXXXX dilution, AS08 330; PsaC 1:XXXXXX dilution, AS10 939; RbcL 1:10000 dilution, AS03 037) and a 1:15000 dilution of the secondary antibody, goat anti-rabbit HRP conjugated (Bio-Rad, Hercules, CA, USA; SKU 1706515, lot 64230791).  Blots were incubated in Amersham ECL Select chemiluminescent substrate (Danaher, Washington, DC, USA) and imaged in a Bio-Rad VersaDoc imager (Bio-Rad, Hercules, CA, USA). Images were quantitated using the ImageJ2 software (version 1.53c).

# Results and discussion {.unnumbered}

@@

## Discussion PAR and growth trends {.unnumbered}

The purpose of this study was to determine the interactive effects of
photoperiod, PAR level, spectral bandwidth and dissolved [O~2~] regimes
on the growth of *P. marinus*, under simulated present day and
hypothetical future ecological niches.

### Photoperiod {.unnumbered}

The representative of HLI clade, *P. marinus* MED4, was completely
excluded from growth under the 4 h photoperiod, and then showed
increasing growth rate with increasing photoperiod. Thus, MED4 can
potentially exploit longer summer photoperiods in temperate zones, but
will remain excluded from growth in winter through its requirement for
more than 4 h of light per day.

The LLIII clade, *P. marinus* SS120, grew under photoperiods longer than
4 h and showed increasing growth rate with increasing photoperiods.
Therefore, SS120 can potentially exploit the longer summer photoperiods
in temperate zones but cannot sustain growth in winter through its
requirement for more than 4 h of light per day.

@@

### PAR Levels {.unnumbered}

The representative of HLI clade, *P. marinus* MED4, was able to grow
under all PAR levels, so long as the photoperiod was longer than 4 h,
although growth rate were slower under the lowest PAR level of 30 µmol
photons m^-2^ s^-1^. Considering the range of PAR levels that MED4 can
utilize, it is reasonable to suggest that MED4 can inhabit not just
shallow depths, where light levels are high, but also deeper regions
characterized by lower light levels, subject to the limitation of a
photoperiod of at least 4 h even after depth attenuation. The photo
regimes of winter temperate zones, due to shorter photoperiods, exclude
MED4 from growth at all depths, however temperate photoperiods and light
levels for the remainder of the year are adequate to support growth.

@@

The representative of HLI clade, *P. marinus* MED4, grew under all
spectral wavebands, but grew slower under green light. Growth rates were
generally comparable under blue, white LED and red lights when exposed
to PAR levels greater than 30 µmol photons m^-2^ s^-1^ for each
photoperiod. At the lower PAR of 30 µmol photons m^-2^ s^-1^, MED4 grew
better under blue light. All growth rates for each spectral waveband
increased with increasing photoperiod. The ability to utilize all
spectral wavebands at high PAR levels and longer photoperiods make the
near surface temperate zones available niches for MED4. MED4 ability to
exploit blue light under low PAR levels and variable photoperiods
greater than 4 h allow MED4 to exploit deeper regions in these temperate
zones, where blue light prevails.

The LLIII clade, *P. marinus* SS120 and the LLIV clade, *P. marinus*
MIT9313, exhibited different growth responses to spectral wavebands
under varying O~2~ conditions, as such, this section will address the
growth response under 250 µM O~2~ experiments, while the interactions of
spectral wavebands under all O~2~ treatments will be discussed in the
next section. SS120, grew under all spectral wavebands at a PAR of 30
µmol photons m^-2^ s^-1^. Growth was completely inhibited under blue
light at PAR levels of 90 and 180 µmol photons m^-2^ s^-1^, but growth
occurred at PAR of 90 µmol photons m^-2^ s^-1^ under all other spectral
wavebands. Generally, growth was strongly inhibited under PAR of 180
µmol photons m^-2^ s^-1^ for all spectral wavebands except for green 16
h photoperiod. The ability of SS120 to grow under low blue light and
longer photoperiods, makes the deep ocean waters of the temperate zone a
viable potential habitat. SS120 is constrained to deeper ocean waters
through its intolerance of higher blue PAR levels.

The LLIV clade, *P. marinus* MIT9313, grew under all spectral wavebands
when exposed to PAR of 30 µmol photons m^-2^ s^-1^ and 90 µmol photons
m^-2^ s^-1^ and was even able to grow under PAR of 180 µmol photons
m^-2^ s^-1^ for all spectral wavebands except blue light. MIT9313 can
thus exploit blue light in deep waters of temperate zones and can even
tolerate moderate depths to exploit higher blue light levels, as long as
the photoperiod is shorter than 12 h.

@@

The representative of HLI clade, *P. marinus* MED4, was able to grow
under both 250 µM and 25 µM O~2~ treatments but was unable to grow under
the lowest O~2~ level of 2.5 µM. MED4 generally exhibited similar growth
response trends and growth rate under blue and red light treatments
under both 25 µM and 250 µM of O~2~. This demonstrated ability to grow
under 25 µM O~2~, coupled with its ability to grow under lower PAR
levels, enables MED4 to potentially inhabit deeper regions where [O~2~]
may vary and low blue light is present, as long as the photoperiod
exceeds 4 h. Our findings are supported by the analysis of TARA Oceans
Project data conducted by @delmontLinkingPangenomesMetagenomes2018 who
reported the presence of *P. marinus* MED4-like genomes at depths
ranging from 5 m to 90 m, representing high to low blue light levels, in
the Pacific South East Ocean. @delmontLinkingPangenomesMetagenomes2018
did not analyze data from depths beyond the subsurface chlorophyll
maximum layer, the layer in the water column where the chlorophyll a
concentration peaks, nor did they report the [O~2~] at those depths but
we have shown, in this study, that MED4 can potentially grow in these
regions. Our findings were also consistent with
@mattisonPhotosyntheticStrategiesWild2020 who analyzed the OceansMap
Protein Portal (OPP) [@saitoDevelopmentOceanProtein2021] for proteins
expressed by vertically stratified *P. marinus* strains at varying
depths and [O~2~], in the ocean minimal zone of the tropical North
Pacific Ocean, and found PSII proteins annotated as MED4 expressed at
depths up to 200 meters with O~2~ of 15 µM.

The LLIII clade, *P. marinus* SS120, was able to grow under both 250 µM
and 25 µM O~2~ treatments but only grew under 2.5 µM O~2~ with a red
light photoperiod of 12 h. The growth responses under 250 µM and 25 µM
O~2~ were very different under the higher PAR level treatments. As
discussed in the spectral waveband section, SS120 did not grow when
treated with 250 µM O~2~ under PAR of 90 and 180 µmol photons m^-2^
s^-1^ blue light. However, SS120 did grow under the lower O~2~ of 25 µM,
photoperiods of 4 and 8 h and PAR of 90 µmol photons m^-2^ s^-1^ blue
light and photoperiods of 8 h and PAR of 180 µmol photons m^-2^ s^-1^
blue light. Under 25 µM O~2~, SS120 grew under all tested white LED
treatments however, did not grow under PAR of 30 µmol photons m^-2^
s^-1^ and 4 h red light photoperiod nor under photoperiods longer than 8
h when treated with PAR of 180 µmol photons m^-2^ s^-1^ red light.
Beyond photoregimes expected of the deep ocean temperate zone waters,
characterized by low levels of blue light and extended photoperiods,
SS120 can also exploit regions closer to the surface where blue light
PAR levels are higher, provided that the [O~2~] are low, as in the OMZ.
Our laboratory results support field observations reported by
@lavinNovelLineagesProchlorococcus2010 who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length
polymorphism analyses, at depths above 40 m where light levels are
higher in the OMZ. @mattisonPhotosyntheticStrategiesWild2020 found PSII
protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in the oxygen minimal zone of the
tropical North Pacific Ocean.

The LLIV clade, *P. marinus* MIT9313, showed enhanced growth responses
under the 25 µM O~2~ treatments for all tested spectral wavebands and
even grew under the highest PAR level of 180 µmol photons m^-2^ s^-1^,
in contrast to growth inhibition under 250 µM O~2~ and 180 µmol photons
m^-2^ s^-1^. Most impressively, MIT9313 has the capacity to grow just as
well, and in some conditions, even better, under high PAR levels of 90
and 180 µmol photons m^-2^ s^-1^, as compared to the lowest PAR of 30
µmol photons m^-2^ s^-1^, when treated with the lower O~2~ of 25 µM, for
all spectral wavebands tested. Interestingly, while generally the growth
rate were slower under the lowest imposed O~2~ of 2.5 µM compared to
O~2~ of 25 µM, MIT9313 was able to grow under all blue light with
photoperiods up to 12 h, along with growth under 2.5 µM O~2~ and some
combinations of red and green light, across photoperiods. Given that
MIT9313 can grow under all photoperiods at the lowest blue light PAR of
30 µmol photons m^-2^ s^-1^, all seasons in the temperate zones are
potentially viable niches. Like SS120, MIT9313 has the capability to
exploit different ocean water layers through its tolerance of higher
light levels under lower [O~2~]. MIT9313 has a remarkable ability to
thrive under very low [O~2~], potentially allowing it to expand into
broader ecological niches. These results support the findings of
@mattisonPhotosyntheticStrategiesWild2020 where PSI protein subunits
annotated as derived from MIT9313 were detected at depths \> 120 m,
along with PSII subunits at depths from 50 m to 200 m in regions where
O~2~ was 15 µM.

Under the 250 µM O~2~ growth experiments and under some photoperiods,
when subjected to the highest PAR of 180 µmol photons m^-2^ s^-1^,
MIT9313 grew when treated under white LED, red and green light and SS120
grew under white LED and green light treatments. However, neither of
these strains exhibited any growth under blue light conditions. Growth
studies have demonstrated that LL clades experience growth inhibition at
PAR levels between 150 and 200 µmol photons m^-2^ s^-1^
[@moorePhotophysiologyMarineCyanobacterium1999]. In the subsequent
chapter, we explore the physiological factors contributing to these
divergent growth responses to different spectral wavebands.

@@

### Results PUR and growth trends {.unnumbered}

Cumulative diel absorbed PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of absorbed light. We then plot observed growth rate vs. cumulative diel absorbed PUR to determine whether growth is a simple response to absorbed light per day across imposed photoperiods and spectral wavebands, or whether photoperiod or spectral wavebands have specific or interactive influences on growth of each ecotype. Cumulative daily absorbed PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed absorbed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h).

Due to the highly effective absorption of pigments in the blue spectral waveband range for all ecotypes, the maximum imposed cumulative diel absorbed PUR under blue light treatment is almost 3 times that of white LED light and about 6 times the red light treatment, despite being derived from the same imposed photoperiods and PUR regimes. As such, only growths rates treated under blue light extend beyond absorbed PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^. A full list of cumulative diel absorbed PUR (µmol photons m^-2^ d^-1^) for all imposed spectral wavebands, PAR and photoperiod regimes can be found in Table \@ref(tab:DailyPURTable).

#### MED4 {.unnumbered}

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative die
equivalent to other photoperiod treatments.
In parallel MED4 showed no growth under 2.5 µM O~2~, under the same level of diel cumulative absorbed PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 showed a simple saturating response of growth to increasing cumulative diel absorbed PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:AllStrainPurFitsPlots)A).

#### SS120 {.unnumbered}

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. A majority of the 4 h photoperiod treatments did not grow under 250 µM O~2~ experiments despite having the same absorbed PUR dose as other experimental treatments. SS120 did not grow when exposed to more than about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of absorbed PUR under any spectral waveband or photoperiod combination under 250 µM O~2~ experiments(Figure \@ref(fig:AllStrainPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel absorbed PUR for cultures growing under 25 µM O~2~.

#### MIT9313  {.unnumbered}

The LLIV clade, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of absorbed PUR, with higher growth rate at over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, flatter response to cumulative diel absorbed PUR under 2.5 µM O~2~. Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of absorbed PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands.  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel absorbed PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ (Figure \@ref(fig:AllStrainPurFitsPlots)C).

@@

The absorption maxima of Chl a~2~ and Chl b~2~ are 442 nm and 478 nm, respectively [@goerickePigmentsProchlorococcusMarinus1992]. At these spectral wavelengths, *P. marinus* efficiently harvests the light energy present in the blue and green waveband range, 400 nm to 525 nm [@morelAvailableUsableStored1978] that penetrates the deep ocean waters. The high Chl a~2~/Chl b~2~ ratio we observed for MED4 is characteristic of the HL clades and the lower Chl a~2~/Chl b~2~ ratios we observed for SS120 and MIT9313 are typical of the LL clades [@moorePhotophysiologyMarineCyanobacterium1999]. The lower amounts of  absorbed PUR in the white LED and green waveband regions for MED4, in comparison to SS120 and MIT9313, are the result of higher Chl a~2~/Chl b~2~ content. Subsequently, calculating absorbed PUR allows direct and meaningful comparison of the energy absorbed by cells on a per photon basis across different wavebands.

In the previous chapter, under 250 µM O~2~ treatments and all PAR of 90 and 180 µmol photons m^-2^ s^-1^ blue light treatments, *P. marinus* SS120, exhibited growth inhibition but was able to exploit all other spectral wavebands under some photoperiods at these PAR levels. Similarly, *P. marinus* MIT9313, showed complete growth inhibition under all PAR 180 µmol photons m^-2^ s^-1^ blue light treatments and only grew under a 90 µmol photons m^-2^ s^-1^, 4 and 8 h blue light photoperiods, yet was able to grow under all other spectral wavebands to varying degrees at these PAR levels. This inhibition initially suggested that blue wavebands triggered photoinhibition at significantly lower levels of Photosynthetically Active Radiation (PAR) compared to other spectral wavebands. However, considering the equivalent cumulative diel absorbed PUR of the blue waveband, all treatments including blue, white LED, red and green waveband exhibited approximately the same growth saturation.

@@

>>>>

Protein samples were thawed and placed in 2 mL FastPrep®-24 bead lysing matrix D vials (MP Biomedicals, Irvine, CA, USA; SKU 116913050) with 200 uL of 1X protein extraction buffer (0.0281 mol L^−1 TRIS base buffer, 0.0211 mol L^−1 TRIS HCl buffer, 0.0735 mol L^−1 lithium dodecyl sulfate, 1.08 mol L^−1 glycerol, 0.5 mmol L^−1 ethylenediaminetetraacetic acid, 0.1 mg·mL−1 PefaBloc SC (AEBSF) protease inhibitor) [@brownFluxCapacitiesAcclimation2008]  XXX note we used a low tris PEB recipe not sure where it came from XXX. Cells were lysed for 3 cycles of 60 seconds at 6.5m s-1 with 60 second intervals on ice between cycles.  The lysing vials were centrifuged at 16100g for 5 minutes, the supernatant was transferred to a 600 uL microcentrifuge tube and centrifuged again at 16100g for 3 minutes.  40 µL of the supernatant were subsampled into labeled vials and stored at -80 °C until analysis. The bicinchoninic acid (BCA) assay was used for total protein determination using a calibration curve of bovine gamma globulin (BGG; Bio-Rad, Hercules, CA, USA) standard ranging from 0 to 1.0 mg mL^-1 range.

All samples and appropriate standards were prepared as described in [@brownFluxCapacitiesAcclimation2008]. Between 0.1 to 4.0 µg of total protein was loaded onto 15 well Invitrogen Bolt 4-12% acrylamide Bis-Tris gels (ThermoFisher Scientific, Waltham, MA, USA). All gels were run for 25 minutes at 200V using Invitrogen Bolt Mini Gel tanks and Invitrogen Bolt MES SDS running buffer (ThermoFisher Scientific, Waltham, MA, USA) and transferred for 60 minutes at 20V onto a PVDF membrane (Bio-Rad, Hercules, CA, USA; SKU 1620177).  Blots were blocked overnight in 2% ECL blocking solution as described by @brownFluxCapacitiesAcclimation2008. The following morning, each blot was incubated in their appropriate antibody obtained from Agrisera (Vännäs, Sweden; FtsH  1:5,000 dilution, AS11 1789; PsbA  1:10000 dilution, AS05 084; PetC 1:XXXXX dilution, ASXXXXX; PsaC 1:XXXXXX dilution, ASXXXXXXX ; RbcL 1:10000 dilution, AS03 037) and a 1:15000 dilution of the secondary antibody, goat anti-rabbit HRP conjugated (Bio-Rad, Hercules, CA, USA; SKU 1706515, lot 64230791).  Blots were incubated in Amersham ECL Select chemiluminescent substrate (Danaher, Washington, DC, USA) and imaged in a Bio-Rad VersaDoc imager (Bio-Rad, Hercules, CA, USA). Images were quantitated using the ImageJ2 software (version 1.53c).

# Results and Discussion {.unnumbered}

@@

### Growth Responses to Photoperiod {.unnumbered}

The representative of HLI clade, *P. marinus* MED4, was completely
excluded from growth under the 4 h photoperiod, andshowed
increasing growth rates with increasing photoperiod. Thus, MED4 can
potentially exploit longer summer photoperiods in temperate zones, but
will likely remain excluded from growth in winter through its requirement for
more than 4 h of light per day, even if water temperature warms into the Clade HLI tolerance range.

The LLIII clade, *P. marinus* SS120, grew under photoperiods longer than
4 h and showed increasing growth rate with increasing photoperiods.
Therefore, SS120 can potentially exploit the longer summer photoperiods
in temperate zones but again cannot sustain growth in winter through its
requirement for more than 4 h of light per day, even if water temperature warms into the Clade LLIII tolerance range.

@@

### Growth Responses to Peak PAR Responses {.unnumbered}

The representative of HLI clade, *P. marinus* MED4, was able to grow
under all PAR levels, so long as the photoperiod was longer than 4 h,
although growth rates were slow under the lowest PAR level of 30 µmol
photons m^-2^ s^-1^. Considering the range of PAR levels that MED4 can
utilize, it is reasonable to suggest that MED4 can inhabit not just
shallow depths, where light levels are high, but also deeper regions
characterized by lower light levels, subject to the limitation of a
photoperiod of at least 4 h even after depth attenuation of light. The photo
regimes of winter temperate zones, due to shorter photoperiods, exclude
MED4 from growth at all depths, however temperate photoperiods and light
levels for the remainder of the year are adequate to support growth.

@@

The representative of HLI clade, *P. marinus* MED4, grew under all
tested spectral wavebands. Growth rates were
generally comparable under blue, white LED and red lights when exposed
to PAR levels greater than 30 µmol photons m^-2^ s^-1^ for each
photoperiod. At the lower PAR of 30 µmol photons m^-2^ s^-1^, MED4 grew
better under blue light. All growth rates for each spectral waveband
increased with increasing photoperiod. The ability to utilize all
spectral wavebands at high PAR levels and longer photoperiods make the
near surface temperate zones available niches for MED4. MED4 ability to
exploit blue light under low PAR levels and variable photoperiods
greater than 4 h allow MED4 to exploit deeper regions in these temperate
zones, where blue light prevails.

The LLIII clade, *P. marinus* SS120 and the LLIV clade, *P. marinus*
MIT9313, exhibited different growth responses to spectral wavebands
under varying O~2~ conditions, as such, this section will address the
growth response under 250 µM O~2~ experiments, while the interactions of
spectral wavebands under all O~2~ treatments will be discussed in the
next section. SS120, grew under all spectral wavebands at a PAR of 30
µmol photons m^-2^ s^-1^. Growth was completely inhibited under blue
light at PAR levels of 90 and 180 µmol photons m^-2^ s^-1^, but growth
occurred at PAR of 90 µmol photons m^-2^ s^-1^ under all other spectral
wavebands. Generally, growth was strongly inhibited under PAR of 180
µmol photons m^-2^ s^-1^ for all spectral wavebands. The ability of SS120 to grow under low blue light and longer photoperiods, makes the deep ocean waters of the temperate zone a viable potential habitat. Under 250 µM O~2~ SS120 is constrained to deeper ocean waters through its intolerance of higher blue PAR levels.

The LLIV clade, *P. marinus* MIT9313, grew under all spectral wavebands
when exposed to PAR of 30 µmol photons m^-2^ s^-1^ and 90 µmol photons
m^-2^ s^-1^ and was even able to grow under PAR of 180 µmol photons
m^-2^ s^-1^ for all spectral wavebands except blue light. MIT9313 can
thus exploit blue light in deeper waters of temperate zones and can even
tolerate moderate depths to exploit higher blue light levels, as long as
the photoperiod is shorter than 12 h.

@@

The representative of HLI clade, *P. marinus* MED4, generally exhibited similar growth
response trends and growth rates under blue and red light treatments
under both 25 µM and 250 µM of O~2~,  but was unable to grow under
the lowest O~2~ level of 2.5 µM. MED4  This demonstrated ability to grow
under 25 µM O~2~, coupled with its ability to grow under lower PAR
levels, enables MED4 to potentially inhabit deeper regions where [O~2~]
may vary and low blue light is present, as long as the photoperiod
exceeds 4 h. Our findings are supported by the analysis of TARA Oceans
Project data conducted by @delmontLinkingPangenomesMetagenomes2018 who
reported the presence of *P. marinus* MED4-like genomes at depths
ranging from 5 m to 90 m, representing high to low blue light levels, in
the Pacific South East Ocean. @delmontLinkingPangenomesMetagenomes2018
did not analyze data from depths beyond the subsurface chlorophyll
maximum layer, the layer in the water column where the chlorophyll a
concentration peaks, nor did they report the [O~2~] at those depths but
we have shown, in this study, that MED4 can potentially grow in these
regions. Our findings were also consistent with
FIGURE REFERENCE who analyzed the OceansMap
Protein Portal (OPP) [@saitoDevelopmentOceanProtein2021] for proteins
expressed by vertically stratified *P. marinus* strains at varying
depths and [O~2~], in the ocean minimal zone of the tropical North
Pacific Ocean, and found PSII proteins annotated as MED4 expressed at
depths up to 200 meters with O~2~ of 15 µM.

The LLIII clade, *P. marinus* SS120 grew under both 250 µM
and 25 µM O~2~ treatments. The growth responses under 250 µM and 25 µM
O~2~ were very different under the higher PAR level treatments. As
discussed in the spectral waveband section, SS120 did not grow when
treated with 250 µM O~2~ under PAR of 90 and 180 µmol photons m^-2^
s^-1^ blue light. However, SS120 did grow under the lower O~2~ of 25 µM,
photoperiods of 4 and 8 h and PAR of 90 µmol photons m^-2^ s^-1^ blue
light and photoperiods of 8 h and PAR of 180 µmol photons m^-2^ s^-1^
blue light. Under 25 µM O~2~, SS120 grew under all tested white LED
treatments however, did not grow under PAR of 30 µmol photons m^-2^
s^-1^ and 4 h red light photoperiod nor under photoperiods longer than 8
h when treated with PAR of 180 µmol photons m^-2^ s^-1^ red light.
Beyond photoregimes expected of the deep ocean temperate zone waters,
characterized by low levels of blue light and extended photoperiods,
SS120 can also exploit regions closer to the surface where blue light
PAR levels are higher, provided that the [O~2~] are low, as in the OMZ.
Our laboratory results support field observations reported by
@lavinNovelLineagesProchlorococcus2010 who found evidence of LLII/III
and LLIV ecotypes, using terminal restriction fragment length
polymorphism analyses, at depths above 40 m where light levels are
higher in the OMZ. @mattisonPhotosyntheticStrategiesWild2020 found PSII
protein subunits annotated as derived from SS120 at all depths ranging
from 20 to 200 m and all [O~2~] in the oxygen minimal zone of the
tropical North Pacific Ocean.

The LLIV clade, *P. marinus* MIT9313, showed enhanced growth responses
under the 25 µM O~2~ treatments for all tested spectral wavebands and
even grew under the highest peak PAR level of 180 µmol photons m^-2^ s^-1^,
in contrast to growth inhibition under 250 µM O~2~ and 180 µmol photons
m^-2^ s^-1^. Most impressively, MIT9313 has the capacity to grow just as
well, and in some conditions, even better, under high PAR levels of 90
and 180 µmol photons m^-2^ s^-1^, as compared to the lowest PAR of 30
µmol photons m^-2^ s^-1^, when treated with the lower O~2~ of 25 µM, for
all spectral wavebands tested. Interestingly, while generally the growth
rate were slower under the lowest imposed O~2~ of 2.5 µM compared to
O~2~ of 25 µM, MIT9313 was able to grow under all blue light with
photoperiods up to 12 h, along with growth under some
combinations of red peak PAR, across photoperiods. Given that
MIT9313 can grow under all photoperiods at the lowest blue light PAR of
30 µmol photons m^-2^ s^-1^, all seasons in the temperate zones are
potentially viable niches, if temperature rises into the Clade LLIV tolerance range. Like SS120, MIT9313 has the capability to exploit different ocean water layers through its tolerance of higher light levels under lower [O~2~]. MIT9313 has a remarkable ability to thrive under very low [O~2~], potentially allowing it to expand into
broader ecological niches. These results support 
FIGUREXXXX where PSI protein subunits
annotated as derived from MIT9313 were detected at depths \> 120 m,
along with PSII subunits at depths from 50 m to 200 m in regions where
O~2~ was 15 µM.

Under the 250 µM O~2~ growth experiments and under some photoperiods,
when subjected to the highest PAR of 180 µmol photons m^-2^ s^-1^,
MIT9313 grew when treated under white LED and red light and SS120
grew under white LED light. However, 
these strains were excluded from growth under high blue peak PAR. Growth
studies have demonstrated that LL clades experience growth inhibition at
PAR levels between 150 and 200 µmol photons m^-2^ s^-1^
[@moorePhotophysiologyMarineCyanobacterium1999].

@@

## PUR and growth responses {.unnumbered}

Cumulative diel absorbed PUR can potentially collapse photoperiod, PAR and spectral wavebands to a common metric of absorbed light. We plot observed growth rate vs. cumulative diel absorbed PUR to determine whether growth is a simple response to absorbed light per day across imposed photoperiods and spectral wavebands, or whether photoperiod or spectral wavebands have specific or interactive influences on growth of each ecotype. Cumulative daily absorbed PUR dose (µmol photons m^-2^ d^-1^) was calculated from the imposed absorbed PUR (µmol photons m^-2^ s^-1^) and photoperiod (h).

Due to the highly effective absorption of pigments in the blue spectral waveband range for all ecotypes, the maximum imposed cumulative diel absorbed PUR under blue light treatment is almost 3 times that of white LED light and about 6 times the red light treatment, despite being derived from the same imposed photoperiods and PUR regimes. As such, only growths rates under blue light extend beyond absorbed PUR of ~ 2 x 10^6^ µmol photons m^-2^ d^-1^.

#### MED4 Clade HLI {.unnumbered}

The representative of HLI clade, *P. marinus* MED4, showed no growth under any 4 h photoperiod treatments, even when a 4 h photoperiod delivered cumulative diel PUR
equivalent to other photoperiod treatments. In parallel MED4 showed no growth under 2.5 µM O~2~, no matter the level of diel cumulative absorbed PUR. 
In contrast, under 250 or 25  µM O~2~, and under any photoperiod greater than 4 h, MED4 growth was well described by a simple saturating response of growth to increasing cumulative diel absorbed PUR, with saturation of growth rate achieved around  1.0 x 10^6^ µmol m^-2^d^-1^ (Figure \@ref(fig:AllStrainPurFitsPlots)A).   XXXXDISTINCT PHOTOPERIOD FITSXXXXX

#### SS120 Clade LLIII {.unnumbered}

The representative of the LLIII clade, *P. marinus* SS120 showed almost no growth under 2.5 µM O~2~ experiments. A majority of the 4 h photoperiod treatments did not grow under 250 µM O~2~ experiments even when a 4 h photoperiod delivered the cumulative diel PUR equivalent to other as other photoperiod treatments. SS120 did not grow when exposed to more than ~1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative diel PUR under any spectral waveband or photoperiod combination, under 250 µM O~2~ experiments(Figure \@ref(fig:AllStrainPurFitsPlots)B).

Under both 25 or 250 µM O~2~ experiments, SS120 growth plateaued by about 5.0 x 10^5^ µmol photons m^-2^ d^-1^ diel PUR, with some scatter among photoperiod and spectral waveband regimes. The onset of growth inhibition extended to higher cumulative diel absorbed PUR for cultures growing under 25 µM O~2~. XXXXDISTINCT PHOTOPERIOD FITSXXXXX

#### MIT9313 Clade LLIV {.unnumbered}

The LLIV clade, *P. marinus* MIT9313, showed growth rising to a plateau by about 5 x 10^5^ µmol photons m^-2^ d^-1^ of cumulative diel PUR, with higher growth rates over a narrower plateau under 25 and 250 µM O~2~, compared to a wider, flatter response to cumulative diel  PUR under 2.5 µM O~2~. Above about 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of cumulative PUR under 250 µM O~2~, MIT9313 showed full inhibition of growth, across photoperiods, and spectral wavebands.  In contrast, under 25 µM O~2~, MIT9313 showed a greatly extended exploitation of higher cumulative diel PUR, with full growth inhibition only above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^.  Similarly, under 2.5 µM O~2~, MIT9313 grew more slowly, but only showed full growth inhibition above about 3.5 x 10^6^ µmol photons m^-2^ d^-1^ cumulative diel PUR (Figure \@ref(fig:AllStrainPurFitsPlots)C).  XXXXDISTINCT PHOTOPERIOD FITSXXXXX

@@

The absorption maxima of Chl a~2~ and Chl b~2~ are 442 nm and 478 nm, respectively [@goerickePigmentsProchlorococcusMarinus1992]. At these spectral wavelengths, *P. marinus* efficiently harvests the light energy present in the blue and green waveband range, 400 nm to 525 nm [@morelAvailableUsableStored1978] that penetrates the deep ocean waters. The high Chl a~2~/Chl b~2~ ratio we observed for MED4 is characteristic of the HL clades and the lower Chl a~2~/Chl b~2~ ratios we observed for SS120 and MIT9313 are typical of the LL clades [@moorePhotophysiologyMarineCyanobacterium1999]. The lower PUR for the white LED waveband for MED4, in comparison to SS120 and MIT9313, are the result of higher Chl a~2~/Chl b~2~ content. Subsequently, estimating  PUR allows direct comparison of the energy absorbed by cells on a per photon basis across different wavebands.

Under 250 µM O~2~ treatments and all PAR of 90 and 180 µmol photons m^-2^ s^-1^ blue light treatments, *P. marinus* SS120, exhibited growth inhibition but was able to exploit all other spectral wavebands under some photoperiods at these PAR levels. Similarly, *P. marinus* MIT9313, showed complete growth inhibition under all PAR 180 µmol photons m^-2^ s^-1^ blue light treatments and only grew under a 90 µmol photons m^-2^ s^-1^, 4 and 8 h blue light photoperiods, yet was able to grow under all other spectral wavebands to varying degrees at these PAR levels. This inhibition initially suggested that blue wavebands triggered photoinhibition at significantly lower levels of Photosynthetically Active Radiation (PAR) compared to other spectral wavebands. However, considering the equivalent cumulative diel absorbed PUR of the blue waveband, all treatments including blue, white LED, red and green waveband exhibited approximately the same growth saturation.

@@

We determined the Photosynthetic Usable Radiation (PUR), the portion of light radiation absorbed differentially by the pigments of the photosynthetic organism [@morelAvailableUsableStored1978] and expressed growth response in terms of cumulative diel PUR.  
We found that approximately 1.0 x 10^6^ µmol photons m^-2^ d^-1^ of PUR is the maximum dosage that the representative LL clades, SS120 and MIT9313 can tolerate under 25 µM O~2~, in contrast to the representative HL clade, MED4 which exhibited growth saturation at the same cumulative diel PUR of 1.0 x 10^6^ µmol photons m^-2^ d^-1^.  Growth under lower O~2~ allowed MIT9313 to substantially increase its exploitation of higher diel PUR.
